### 2021-08-27 paper/post_tci_supplementary.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\bibliographystyle{IEEEtran}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{geometry}
\usepackage[colorlinks=true]{hyperref}
\title{Supplemental material for `Eliminating object prior-bias from sparse-projection
tomographic reconstructions'}
\begin{document}
\maketitle
%This supplementary material consists of details of the optimization routine used in our method, links to view our 3D reconstructions and results of 2D reconstructions for different projection views.

\tableofcontents
\newpage

%--------------------------------Links to videos-------------------------------------------------
\newpage
\section{Links to view our 3D reconstructions}
The 3D reconstructions, corresponding to the images shown in Figs.~14,~16 and 19 in the main paper, can be better visualized in the form of a video. Hence, please refer to the following videos to view the 3D datasets and reconstruction results:

\begin{itemize}
\item \textbf{Okra:} \\

  \url{https://www.dropbox.com/sh/x3yjfzsc5633d1t/AAAaVOXvTbc06a2DomPO1kSna?dl=0}\\
\item \textbf{Potato:} \\

  \url{https://www.dropbox.com/sh/6mja5uh8fa2h0fc/AAA2h4VsLxAiXcEIhmTlM1xha?dl=0}\\
\item \textbf{Sprouts:}\\

  \url{https://www.dropbox.com/sh/qtgeyckvbxiaz3s/AADhR2zRUUN4qXSXraErcqSea?dl=0}\\
\end{itemize}

\newpage

\end{document}