### 2021-09-20 paper/changes.tex
updated changes with some reformatting and updated post_tci in two places

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\bibliographystyle{IEEEtran}
\usepackage{caption}
\usepackage{subcaption, tabularx}
\usepackage{xcolor}
\usepackage{boxedminipage}
\usepackage{a4wide}
\usepackage{subcaption}
\usepackage[colorlinks=true]{hyperref}
\title{Response to Reviewers and Associate Editor}
\begin{document}
\date{}
\maketitle

This paper was submitted earlier, and out of the three reviewers, two
of them had pronounced this as ``publish unaltered'' and one reviewer
(R1) did not recommend resubmission. The handling Editor has
encouraged a resubmission of this paper taking into account the
comments of R1.  We thank all reviewers for their remarks and in this
document, we explain the steps we have taken to address the reviews
received in our previous submission to TCI. R1 makes the following
comments, and we quote excerpts in blue here:
\begin{quote}
  \textcolor{blue}{The authors have spent considerable effort in revising the
  manuscript and addressing the concerns of the reviewers. I would
  like to thank the authors for their effort. I believe the manuscript
  has improved because of these revisions, especially in the structure
  of the paper and the discussion of related work. Unfortunately, some
  of my main concerns with regards to the experiments, the accuracy of
  the proposed method, and the presentation of the findings remain
  present in the revised manuscript. In some cases the revisions have
  even increased my concerns about certain aspects. Mainly, the
  included experiments still do not clearly show that the proposed
  method improves upon existing ‘simple’ prior-based methods
%% , and the
%%   new results even show that there is evidence that FDK is more
%%   accurate than the proposed method exactly in the region that changed
  }
  \end{quote}

The simpler method is based on the well-known ``Total Variation''. 
In particular, we address the following main concern of the reviewer 
\begin{quote}  
    \textcolor{blue}{Note that the total
variation minimization results are much better than the l1-ls results that the authors show in the
paper}
\end{quote}

    by rewriting the paper considerably.  We believe the paper
    addresses this (correct) stinging criticism of R1.  In fact R1 has taken
    pains in showing how the Total-Variation method (and sometimes
    FDK) is better than our earlier Compressed Sensing plus
    spatially-varying prior-based reconstruction, and we gratefully
    acknowledge the same.

Again, this is correct for the dataset shown, and we accept this to be
true.

  We re-wrote our code and upon further analysis observed that TV is
  indeed better suited as a baseline reconstruction method for our
  datasets and our subsampling choices. Hence, our current paper
  presents all results using TV regularization coupled with
  spatially-varying technique, and demonstrates  benefits over
  TV-only and backprojection-only methods.  \emph{We have eliminated our
  previous implicit claim that the CS based scheme is optimal for the
  dataset provided. }

  As far as FDK is concerned, the current results indeed show the
  superiority of the TV-based scheme (as also noted by the reviewer).

    That said, the goal of our previous and curent work is to provide
    a technique that can improve upon a chosen baseline reconstruction
    when measurements are extremely sparse.  One may choose any
    baseline reconstruction (including an inferior -- for this dataset
    -- Compressed sensing prior) for pedagogical, historical, or
    commercial reasons, and given that, our key question is -- is
    there a way to improve?  The spatially-varying method of
    moderating the prior settles this question.

We also take this opportunity to address other comments given by R1
and the handling editor.

 \begin{itemize}
 \item    \textcolor{blue}{...the paper should at least include a (short) discussion about the
     possibility of using deep learning for this, and its disadvantages.}

   We accept this point, and have now discussed the difficulties in
   using deep-learning techniques for the current problem in Sec.~7C
   of our paper. We also describe them here below.

\textcolor{red}{Comment from Sharat:  Since the material is in the text, there is no need to mention it here.}

  In a machine learning based setting, our goal of detecting new
  changes is a `prediction' problem in a continuous solution space
  i.e., \textit{``given intensities of a voxel at various time
    instants in a longitudinal setting and partial measurements of the
    voxel at the current time, what will be the intensity of the same
    voxel at the current time instant?''} This estimation can be
  learnt by a deep neural network if there are hundreds of labeled
  data.  However, generalization of this across multiple datasets
  poses a question mark.

  %% thousands of time instants at which the previous voxel
  %% intensities are known for each voxel. This amount of data is not
  %% available to us, and may be hard to acquire because each specimen
  %% is usually scanned only a couple of times in any longitudinal
  %% study.

  Further, a specific `event-of-interest' may occur just once in a
  longitudinal study and hence training on data of all previous time
  instants may be misleading. If however, we use measurements from
  identical and full-cycle longitudinal studies, including
  events-of-interest, of similar specimen (instead of the `same'
  specimen as we used), a deep-learning technique may be applied. We
  see this as an extension of using object-prior generated from the
  same object, and hence we have not explored this direction.

  Another avenue for using deep-networks \textit{within} our current
  work is to replace the currently used fixed eigenspace provided by
  PCA by a learnt feature basis provided by a network such as an
  autoencoder. For each dataset, an autoencoder may be trained to
  learn a specific set of latent features (this might again require
  atleast a few tens of volumes). We see this as a future direction of
  work.

\item \textcolor{blue}{There are no added comparisons with popular existing methods for
  reconstruction with a small number of projections (TV, algebraic techniques, etcetera)}\\

 Among the family of algebraic techniques, we have compared our method
 with Algebraic Reconstruction Technique (ART) and Simultaneous
 Algebraic Reconstruction Technique (SART). We do not reconstruct
 using Discrete Algebraic Reconstruction Technique (DART) since there
 is no prior information about the number of possible attentuation
 coefficients in the test object. Among other iterative techniques, we
 have used Simultaneous Iterative Reconstruction Technique (SIRT),
 Total Variation (TV) and Compressed Sensing (CS) with Haar wavelet
 and DCT as sparsity basis. The details of solvers used for these
 algorithms and their reconstructions are in Page.~7 (Section-6A) of
 the paper. For CS reconstruction, we observed that Haar wavelet and
 DCT were best suited for our dataset among various other possible
 sparsity basis.

\item \textcolor{blue}{Why did the authors not choose the standard
  settings for SSIM?...I don’t understand the reason given for not
  including RMSE: the values don’t have to be optimized again -- you
  could just optimize for SSIM but at least still report RMSE
  comparisons.}

  The reason for not reporting RMSE stems from the optimization
  process. The optimization routine we use transforms the
  reconstruction to a different space.  That is, in our experiments,
  we observed that for a given dataset and a set of projection
  measurements, the intensity span of histograms of reconstructed
  volumes differ across various methods and solvers. As a result, the
  RMSE values are not particularly useful.  We provide the numbers
  here to show the improvement trend but we believe it does not serve
  the interest of the community to report these.

  For the same reason, the standard setting of SSIM in matlab (for
  instance) emphasizes the intensity. We found that even when the
  quality of reconstruction is acceptable visually, the SSIM values
  were not acceptable.  We choose
  to focus on preserving the structures on our reconstructions, and
  not rely on the absolute intensity values alone. Hence, we choose
  SSIM with a higher weightage to structure-preservation. For our
  record, we have computed RMSE values as well and they are presented
  here in Tables~\ref{table:potato_rmse} and~\ref{table:sprouts_rmse}.

\begin{table}[!h]
  \centering
 \caption{RMSE within the RoI of 3D reconstructed Potato volume from various
    methods.}
\begin{tabular}{|l|c|c|c|c|}
\hline 
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
\\ \hline  0.29 & 0.24 & \textcolor{red}{0.18}
\\ \hline
\end{tabular}
\label{table:potato_rmse}
\end{table}

\begin{table}[!h]
  \centering
 \caption{RMSE within the RoI of 3D reconstructed Sprouts volume from various
    methods.}
\begin{tabular}{|l|c|c|c|c|}
\hline 
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
\\ \hline  0.53 & 0.48 & \textcolor{red}{0.30}
\\ \hline
\end{tabular}
\label{table:sprouts_rmse}
\end{table}

\item \textcolor{blue}{A discussion about computational costs (time,
  memory) should be included in the paper.}

  We now present details of both time and space complexities of our technique in Section.7B of our paper.

Most of the computation time in our technique is used for the
generation of low quality eigen-space by reconstructing each of the
templates using TV and FDK. If the set of projection views for every
test is fixed a priori, the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. With
this assumption, Table~\ref{table:compute_time} shows the computation
times for various stages of reconstruction of a 2D slice of Liver on
MATLAB 2019a on an AMD 2920X 12-Core Processor machine with 64GB
RAM. If however, new sets of projection views for every test are to be
allowed, the low quality reconstructions can still be performed
efficiently using parallelization since these reconstructions are
completely independent. Here below, we describe the time complexity of
spatially-varying prior technique.  If
\begin{itemize}
\item $N =$ number of voxels of the volume
\item $T =$ number of templates available, and
\item $F =$ time taken for iterative TV optimization,
\end{itemize}
then, the computational complexity is sum of the following
\begin{itemize}
\item $O(NT^2+T^3)$ for creating the eigenspace,
\item $O(NT^2)$ for deriving the weights map, and
\item time $F$  taken for iterative minimization
\end{itemize}
where $O(.)$ is the notation for `Big O' used in complexity analysis. 
In terms of space, the complexity is $O(T)$ in order to store the templates and their low quality reconstructions. We only need to store a single weights map volume regardless of the number of templates available.

\begin{table}[]
  \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
Backprojection             & TV                        & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map\end{tabular} & \begin{tabular}[c]{@{}l@{}}Final reconstruction\\ with spatially-varying \\ prior\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.40} & \multicolumn{1}{c|}{8.44} & \multicolumn{1}{c|}{0.04}                                             & \multicolumn{1}{c|}{46.73}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\end{itemize}

\end{document}